A Unified Treatment of Tractability for \\ Approximation Problems Defined on Hilbert Spaces

### 2023-11-15 GeneralizedTractability2023Sept21.tex
Updates from Overleaf

\title{A Unified Treatment of Tractability for Approximation Problems Defined on Hilbert Spaces}

@@

%The non-decreasing ordering of the eigenvalues for each $d$ facilitates the proof of conditions for tractability, but our conditions do not depend on the ordering \peter{[PETER: is that really the case?]}.

We make the technical assumption that $\SOL_d$ is a compact operator. This implies (see, e.g., \cite{NovWoz08a}) that $\lambda_{i,d}$ converges to zero as $i$ tends to infinity for every $d$, which in turn implies that the problem is solvable by suitable algorithms. To avoid trivial cases, we assume that there are an infinite number of positive eigenvalues for every $d$.

@@

Strictly speaking, we are considering the \textit{absolute} error criterion in this paper.
Alternatively, one could also consider the \textit{normalized} error criterion, where the error in \eqref{eq:errcrit} is divided by the so-called \textit{initial error}, which is the error without sampling the function, or, equivalently, the operator norm of $\SOL_d$. In the latter case, it is known (see, e.g., again \cite{NovWoz08a}) that the information complexity is given by
\[
\comp^{\rm norm}(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\, \lambda_{1, d}\}.
\]
This means that, essentially, in the tractability analysis below, all criteria would be normalized by the first \peter{singular value} $\lambda_{1,d}$. For the sake of brevity and clarity of notation, we restrict ourselves to the absolute error criterion here, but presumably analogous results hold for the normalized setting.

@@

We want to bound the information complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$, and identify conditions on the singular values that are easier to verify that are equivalent to this bound on the information complexity holding. We note that generalized tractability specified by a function $T$ has been considered before in the literature. In particular, there are several papers by Gnewuch and Wo\'{z}niakowski (see \cite{GW06,GW08,GW09,GneWoz11a}), and also \cite[Chapter 8]{NovWoz08a} is devoted to this topic. To our understanding, however, these references mainly focus on the special case of tensor product problems, a restriction we do not make.  Moreover, the conditions for tractability derived there focus mostly on relating the properties of the function $T$ to the different varieties of tractability, and not in terms of sums of (functions of) the \peter{singular values} of a problem. Hence, the present paper adds to what is presently known on this subject. 
In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we define our approximation problem to be tractable if $\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$. The parameter $\vp$ is an $s$-dimensional vector with $s \geq 1$, where $\vp \geq \vzero$ implies that each component of $\vp$ is non-negative. In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase sub-exponentially  with increasing $\varepsilon^{-1}$ and/or $d$, however, most of our theorems do not require such an assumption.

\Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} below all take the form
\begin{multline}
  \exists \,C_{\vp} > 0 \text{ with }
    \comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},\thed,\vp) \quad \forall (\varepsilon^{-1},d) \in \Omega \\
    \iff   \exists L_{\vp} \peter{>0} \text{ with }
    \sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,\thed,\vp) \rceil}^{\theUB(d)} \frac{1}{T(\lambda^{-1}_{i,d},\thed,\vp)}< \infty, \\
     \thed \in \{1,d\}, \ \theM(d) := \inf \{\varepsilon : (\varepsilon^{-1},d) \in \Omega \}, \
     \theUB(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \theM(d) \}, &   \theM(d) > 0, \\
        \infty, & \theM(d) = 0.
        \end{cases}
\end{multline}
The differences in these theorems are in the choices of $\thed$ and $\Omega$. Although some of these theorems are special cases of others, we prove each separately to allow the reader to become familiar with the arguments used as additional layers of complexity are added.
\Cref{thm:subhT} has a somewhat different form than above.

@@

The remainder of this paper is organized as follows. In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider strong tractability when there is an upper bound on the information complexity that is independent of the dimension, $d$. We then turn to tractability  when the bound on the information complexity may depend on both $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of tractability functions for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and introduce other notions of tractability. A generalized notion of weak tractability is given in Section \ref{sec:subh}. In Section \ref{sec:restricted}, we consider \peter{tractability on} a restricted domain .

@@

For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation} \label{eq:Tnondecreasing}
	T \text{ is non-decreasing in all variables,}
\end{equation}
which implies that the problem is expected to be no easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$\footnote{This is why we make $T$ a function of $\varepsilon^{-1}$ rather than of $\varepsilon$.} Furthermore, increasing $\vp$ allows for a possibly looser bound on the information complexity.  Since there are an infinite number of positive \peter{singular values}, it also makes sense to assume that
\begin{equation}\label{eq:Tlim}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit, which we assume will always be positive:
\begin{equation}\label{eq:T0}	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp) = \inf_\varepsilon T(\varepsilon^{-1},d,\vp) \ge T(0,1,\vzero) > 0.
\end{equation}
\end{subequations}
Finally, there is a technical assumption required for our analysis in \Cref{sec:spt,sec:restricted}.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp)   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}

@@

\begin{table}
    \caption{Common forms of the tractability function, $T$.}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		%\text{algebraic strong polynomial} & \max\{1,\varepsilon^{-p}\} & 1\\
		\text{algebraic polynomial} & \max\{1,\varepsilon^{-p}\}d^{q} & d^{q}\\
		%\text{exponential strong polynomial} &  [\max\{1,\log(1 + \varepsilon^{-1}\}]^p & 1\\
		\text{exponential polynomial} &
		[\max\{1,\log(1 + \varepsilon^{-1})\}]^p  d^{q} & d^{q} \\
        \text{algebraic quasi-polynomial} &
        \exp\{p(1+\log(\max\{1,\log(1 + \varepsilon^{-1})\})(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{exponential quasi-polynomial} &
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
\label{tab:commonT}
\end{table}

@@

\begin{proof}
\textbf{Sufficient condition:}\newline
We make the first part of the argument in some generality so that it can be reused in the proof of Theorem \ref{thm_main_tract2} for tractability.  Fix any $\varepsilon > 0$ and any $d \in \naturals$, and let $\thed \in \naturals$ be arbitrary. %\delta replaced d'
Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},\thed,\vp)$ are non-decreasing in $i$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},\thed,\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},\thed,\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},\thed,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) } \le \frac{1}{T(\varepsilon^{-1},\thed,\vp)} \\ \forall \thed, n, N \in \naturals.
\end{multline}
We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer \peter{and} $n - N +1$ is assumed to be positive.

@@

\bigskip
\noindent \textbf{Necessary condition:} \\
Suppose that we have strong
$T$-tractability as defined in \eqref{eq:strong_tract_def}. That is, for some $\vp \ge \vzero$ there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)
\qquad \forall \varepsilon > 0, \ d \in \mathbb{N}.
\]
Since the sequence of \peter{singular values} $\lambda_{1,d}, \lambda_{2,d}, \ldots $ is non-increasing, we have
\begin{equation}\label{eq:lambda_K_strong3}
\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1,d}\le \varepsilon \qquad \forall \varepsilon > 0, \ d \in \mathbb{N}.
\end{equation}

@@

\noindent \textbf{Necessary condition:}\\
Suppose that we have
$T$-tractability. That is, for some $\vp \ge \vzero$, there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon > 0,  \ d \in \naturals.
\]
Since the sequence of \peter{singular values} $\lambda_{1,d}, \lambda_{2,d}, \ldots $ is non-increasing, we have
\begin{equation}\label{eq:lambda_K4}
	\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},d,\vp)\rfloor +1,d}\le \varepsilon \qquad \forall \varepsilon > 0,  \ d \in \naturals.
\end{equation}
Define the integers
\[
i (\varepsilon,d,\vp):= \lfloor C_{\vp}\, T(\varepsilon^{-1},d,\vp)\rfloor +1, \qquad
i (\infty,d,\vp) := \lfloor C_{\vp}\, T(0,d,\vp)\rfloor +1 \ge 1.
\]
Thus, it follows by \eqref{eq:lambda_K4} that $\lambda_{i(\varepsilon,d,\vp),d} \le \varepsilon$.
Note furthermore that we always have
\[
i(\varepsilon,d,\vp)\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{i(\varepsilon,d,\vp),d}^{-1},d,\vp)+1 \qquad \forall \varepsilon > 0, \ d \in \naturals,
\]
since
$T(\cdot,d,\vp)$ is non-decreasing.

@@

Note now that, for any $d\in\naturals$,
\begin{align*}
 i(\infty,d,\vp') +1  & = \lfloor C_{\vp'}\, T(0,d,\vp')\rfloor +2 \le  \left[C_{\vp'} + \frac{2}{T(0,d,\vp')}   \right]\, T(0,d,\vp')\\
 & \le  \underbrace{\left[C_{\vp'} + \frac{2}{T(0,1,\vp')}   \right]}_{=:L_{\vp'}}\, T(0,d,\vp')
 \le \lceil L_{\vp'} T(0,d,\vp')\rceil .
\end{align*}
For this choice of $L_{\vp'}$ we have
\[
\sup_{d\in\naturals}\,\, \sum_{i=\lceil \kachi{L_{\vp'}} T(0,d,\vp')\rceil}^\infty \frac{1}{T(\lambda_{i,d}^{-1},d, \vp')}
\le
\sup_{d\in\naturals} \, \, \sum_{i=i (\infty,d,\vp')+1}^\infty \frac{1}{T (\lambda_{i,d}^{-1},d,\vp')} <
\infty.
\]
This yields \eqref{eq:tractiff4} with $\vp$ replaced by $\vp'$, so we see the necessity of $\eqref{eq:tractiff4}$.

@@

Next we show that if $\vp^* \in \widetilde{\cp}_{\textup{trct}}$, then $\vp^* \in \cp_{\textup{trct}}$.  If \kachi{$\vp^* \in \widetilde{\cp}_{\textup{trct}}$}, then \eqref{eq:tractiff4}  holds for all $\vp \in (\vp^*,\vinfty)$.  By the argument to prove the sufficient condition above, it follows that \eqref{eq:tract_def} must also hold for all $\vp \in (\vp^*,\vinfty)$. Thus, $\vp^* \in \cp_{\textup{trct}}$.

@@

\section{Examples} \label{sec:examples}
%In this section, we illustrate Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2} for the types of tractability given in the table in Section \ref{sec:not}.
In this section, we present six examples relating to Theorem \ref{thm_main_strong_tract2} and Theorem \ref{thm_main_tract2}, both of which were \peter{shown} in Section \ref{sec:spt}, for the various notions of tractability that are listed in Table \ref{tab:commonT}. In \peter{the} subsequent sections, we will continue to present various examples.

@@

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2,
\] where in this case $\vp = (p,q)$.
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty \min\{1,\lambda_{i,d}^p\}  < \infty \text{ for some } p,q \ge 0, L_{(p,q)} \peter{>0}.
\]
Furthermore, for optimality, let $\widetilde{\cp}_{\textup{ALG-trct}} : = \{(p^*,q^*) : \text{the above condition holds }\forall (p,q) \in (p^*,\infty)\times(q^*,\infty) \}$. Therefore, the set of optimal $(p,q)$ is the set 
\[
    \{(p^*,q^*) \in \widetilde{\cp}_{\textup{ALG-trct}} : (p^*,q^*) \notin [\widetilde{p},\infty) \times [\widetilde{q},\infty)\, \, \forall (\widetilde{p},\widetilde{q}) \in \widetilde{\cp}_{\textup{ALG-trct}}\setminus \{p^*,q^*\}  \}.
\]
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NovWoz08a}.

Having defined the tractability function \peter{as above}, we will now examine a specific case where we always set $d=1$,i.e.,
\[
T(\varepsilon^{-1},1,\vp)= \max\{1,\varepsilon^{-\vp}\}\qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]
where in this case $\vp=p$ is a scalar.

%\peter{[But this is not what we discussed on Wednesday. The idea is that we use the same tractability function as for algebraic polynomial tractability, but consider the special case where we always set $d=1$,i.e., we use 
%\[
 %T(\varepsilon^{-1},1,\vp)= \max\{1,\varepsilon^{-p}\}\\qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2,\]
%as an upper bound on the information complexity. Then everything would correspond to our definitions of $T$-tractability and strong $T$-tractability. Kachi, can you please make this change here and also in Example 2? So, please make it clear that the tractability function for strong polynomial tractability is a specialization of the tractability function for polynomial tractability.]
%}

>>>>

\title{A Unified Treatment of Tractability for \\ Approximation Problems Defined on Hilbert Spaces}

@@

We make the technical assumption that $\SOL_d$ is a compact operator. This implies (see, e.g., \cite{NovWoz08a}) that $\lambda_{i,d}$ converges to zero as $i$ tends to infinity for every $d$, which in turn implies that the problem is solvable by suitable algorithms. To avoid trivial cases, we assume that there are an infinite number of positive singular values for every $d$.

@@

Strictly speaking, we are considering the \textit{absolute} error criterion in this paper.
Alternatively, one could also consider the \textit{normalized} error criterion, where the error in \eqref{eq:errcrit} is divided by the so-called \textit{initial error}, which is the error without sampling the function, or, equivalently, the operator norm of $\SOL_d$. In the latter case, it is known (see, e.g., again \cite{NovWoz08a}) that the information complexity is given by
\[
\comp^{\rm norm}(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\, \lambda_{1, d}\}.
\]
This means that, essentially, in the tractability analysis below, all criteria would be normalized by the first singular value $\lambda_{1,d}$. For the sake of brevity and clarity of notation, we restrict ourselves to the absolute error criterion here, but presumably analogous results hold for the normalized setting.

@@

We want to bound the information complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$, and identify conditions on the singular values that are easier to verify that are equivalent to this bound on the information complexity holding. We note that generalized tractability specified by a function $T$ has been considered before in the literature. In particular, there are several papers by Gnewuch and Wo\'{z}niakowski (see \cite{GW06,GW08,GW09,GneWoz11a}), and also \cite[Chapter 8]{NovWoz08a} is devoted to this topic. To our understanding, however, these references mainly focus on the special case of tensor product problems, a restriction we do not make.  Moreover, the conditions for tractability derived there focus mostly on relating the properties of the function $T$ to the different varieties of tractability, and not in terms of sums of (functions of) the singular values of a problem. Hence, the present paper adds to what is presently known on this subject. 
In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we define our approximation problem to be tractable if $\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$. The parameter $\vp$ is an $s$-dimensional vector with $s \geq 1$, where $\vp \geq \vzero$ implies that each component of $\vp$ is non-negative. In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase sub-exponentially  with increasing $\varepsilon^{-1}$ and/or $d$, however, most of our theorems do not require such an assumption.

\Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} below all take the following form:  for a given $\thed \in \{1,d\}$ and $\Omega \subseteq (0,\infty) \times \mathbb{N}$,
\begin{multline}
  \exists \,C_{\vp} > 0 \text{ with }
    \comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},\thed,\vp) \quad \forall (\varepsilon^{-1},d) \in \Omega \\
    \iff   \exists L_{\vp} >0 \text{ with }
    \sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,\thed,\vp) \rceil}^{\theUB(d)} \frac{1}{T(\lambda^{-1}_{i,d},\thed,\vp)}< \infty, \\
     \text{where } \theM(d) := \inf \{\varepsilon : (\varepsilon^{-1},d) \in \Omega \}, \
     \theUB(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \theM(d) \}, &   \theM(d) > 0, \\
        \infty, & \theM(d) = 0.
        \end{cases}
\end{multline}
%The differences in these theorems are in the choices of $\thed$ and $\Omega$. 
Although some of these theorems are special cases of others, we prove each separately to allow the reader to become familiar with the arguments used as additional layers of complexity are added.
\Cref{thm:subhT} has a somewhat different form than above.

@@

The remainder of this paper is organized as follows. In Section \ref{sec:spt}, we give the definitions of the considered tractability notions, in particular (strong) tractability. We consider strong tractability when there is an upper bound on the information complexity that is independent of the dimension, $d$. We then turn to tractability  when the bound on the information complexity may depend on both $d$ and $\varepsilon^{-1}$. In Section \ref{sec:examples} we provide examples of tractability functions for both the algebraic and exponential cases. We also consider other notions of tractability such as quasi-polynomial tractability, and introduce other notions of tractability. A generalized notion of weak tractability is given in Section \ref{sec:subh}. In Section \ref{sec:restricted}, we consider tractability on a restricted domain, $\Omega$.

@@

For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation} \label{eq:Tnondecreasing}
	T \text{ is non-decreasing in all variables,}
\end{equation}
which implies that the problem is expected to be no easier by decreasing $\varepsilon$, the tolerance, or increasing $d$, the dimension\footnote{This is why we make $T$ a function of $\varepsilon^{-1}$ rather than of $\varepsilon$.}. Furthermore, increasing $\vp$ allows for a possibly looser bound on the information complexity. Since there are an infinite number of positive singular values, it also makes sense to assume that
\begin{equation}\label{eq:Tlim}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit, which we assume will always be positive:
\begin{equation}\label{eq:T0}	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp) = \inf_\varepsilon T(\varepsilon^{-1},d,\vp) \ge T(0,1,\vzero) > 0.
\end{equation}
\end{subequations}
Finally, there is a technical assumption required for our analysis in \Cref{sec:spt,sec:restricted}.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp)   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in [1,\infty).
\end{equation}

@@

\begin{table}
    \caption{Common forms of the tractability function, $T$.}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		%\text{algebraic strong polynomial} & \max\{1,\varepsilon^{-p}\} & 1\\
		\text{algebraic polynomial} & \max\{1,\varepsilon^{-p}\}d^{q} & d^{q}\\
		%\text{exponential strong polynomial} &  [\max\{1,\log(1 + \varepsilon^{-1}\}]^p & 1\\
		\text{exponential polynomial} &
		[\max\{1,\log(1 + \varepsilon^{-1})\}]^p  d^{q} & d^{q} \\
        \text{algebraic quasi-polynomial} &
        \exp\{p(1+\log(\max\{1, \varepsilon^{-1}\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{exponential quasi-polynomial} &
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
\label{tab:commonT}
\end{table}

@@

\begin{proof}
\textbf{Sufficient condition:}\newline
We make the first part of the argument in some generality so that it can be reused in the proof of Theorem \ref{thm_main_tract2} for tractability.  Fix any $\varepsilon > 0$ and any $d \in \naturals$, and let $\thed \in \naturals$ be arbitrary. %\delta replaced d'
Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},\thed,\vp)$ are non-decreasing in $i$. In particular, for  any $N\in\naturals$ with $n\ge N$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},\thed,\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},\thed,\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - N +1 \ge T(\varepsilon^{-1},\thed,\vp) \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},\thed,\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^\infty \frac{1}{T(\lambda_{i, d}^{-1},\thed,\vp) } \le \frac{1}{T(\varepsilon^{-1},\thed,\vp)} \\ \forall \thed, n, N \in \naturals.
\end{multline}
We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer and $n - N +1$ is assumed to be positive.

@@

\bigskip
\noindent \textbf{Necessary condition:} \\
Suppose that we have strong
$T$-tractability as defined in \eqref{eq:strong_tract_def}. That is, for some $\vp \ge \vzero$ there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)
\qquad \forall \varepsilon > 0, \ d \in \mathbb{N}.
\]
Since the sequence of singular values $\lambda_{1,d}, \lambda_{2,d}, \ldots $ is non-increasing, we have
\begin{equation}\label{eq:lambda_K_strong3}
\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1,d}\le \varepsilon \qquad \forall \varepsilon > 0, \ d \in \mathbb{N}.
\end{equation}

@@

\noindent \textbf{Necessary condition:}\\
Suppose that we have
$T$-tractability. That is, for some $\vp \ge \vzero$, there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon > 0,  \ d \in \naturals.
\]
Since the sequence of singular values $\lambda_{1,d}, \lambda_{2,d}, \ldots $ is non-increasing, we have
\begin{equation}\label{eq:lambda_K4}
	\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},d,\vp)\rfloor +1,d}\le \varepsilon \qquad \forall \varepsilon > 0,  \ d \in \naturals.
\end{equation}
Define the integers
\[
i (\varepsilon,d,\vp):= \lfloor C_{\vp}\, T(\varepsilon^{-1},d,\vp)\rfloor +1, \qquad
i (\infty,d,\vp) := \lfloor C_{\vp}\, T(0,d,\vp)\rfloor +1 \ge 1.
\]
Thus, it follows by \eqref{eq:lambda_K4} that $\lambda_{i(\varepsilon,d,\vp),d} \le \varepsilon$.
Note furthermore that we always have
\[
i(\varepsilon,d,\vp)\le C_{\vp}\, T(\varepsilon^{-1},d,\vp)+1 \le C_{\vp} T(\lambda_{i(\varepsilon,d,\vp),d}^{-1},d,\vp)+1 \qquad \forall \varepsilon > 0, \ d \in \naturals,
\]
since
$T(\cdot,d,\vp)$ is non-decreasing.

@@

Note now that, for any $d\in\naturals$,
\begin{align*}
 i(\infty,d,\vp') +1  & = \lfloor C_{\vp'}\, T(0,d,\vp')\rfloor +2 \le  \left[C_{\vp'} + \frac{2}{T(0,d,\vp')}   \right]\, T(0,d,\vp')\\
 & \le  \underbrace{\left[C_{\vp'} + \frac{2}{T(0,1,\vp')}   \right]}_{=:L_{\vp'}}\, T(0,d,\vp')
 \le \lceil L_{\vp'} T(0,d,\vp')\rceil .
\end{align*}
For this choice of $L_{\vp'}$ we have
\[
\sup_{d\in\naturals}\,\, \sum_{i=\lceil L_{\vp'}T(0,d,\vp')\rceil}^\infty \frac{1}{T(\lambda_{i,d}^{-1},d, \vp')}
\le
\sup_{d\in\naturals} \, \, \sum_{i=i (\infty,d,\vp')+1}^\infty \frac{1}{T (\lambda_{i,d}^{-1},d,\vp')} <
\infty.
\]
This yields \eqref{eq:tractiff4} with $\vp$ replaced by $\vp'$, so we see the necessity of $\eqref{eq:tractiff4}$.

@@

Next we show that if $\vp^* \in \widetilde{\cp}_{\textup{trct}}$, then $\vp^* \in \cp_{\textup{trct}}$.  If $\vp^* \in \widetilde{\cp}_{\textup{trct}}$, then \eqref{eq:tractiff4}  holds for all $\vp \in (\vp^*,\vinfty)$.  By the argument to prove the sufficient condition above, it follows that \eqref{eq:tract_def} must also hold for all $\vp \in (\vp^*,\vinfty)$. Thus, $\vp^* \in \cp_{\textup{trct}}$.

@@

\section{Examples} \label{sec:examples}
%In this section, we illustrate Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2} for the types of tractability given in the table in Section \ref{sec:not}.
In this section, we present six examples relating to Theorem \ref{thm_main_strong_tract2} and Theorem \ref{thm_main_tract2}, both of which were shown in Section \ref{sec:spt}, for the various notions of tractability that are listed in Table \ref{tab:commonT}. In the subsequent sections, we will continue to present various examples.

@@

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2,
\] where in this case $\vp = (p,q)$.
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty \min\{1,\lambda_{i,d}^p\}  < \infty \text{ for some } p,q \ge 0, L_{(p,q)} >0.
\]
Furthermore, for optimality, let $\widetilde{\cp}_{\textup{ALG-trct}} : = \{(p^*,q^*) : \text{the above condition holds }\forall (p,q) \in (p^*,\infty)\times(q^*,\infty) \}$. Therefore, the set of optimal $(p,q)$ is the set 
\[
    \{(p^*,q^*) \in \widetilde{\cp}_{\textup{ALG-trct}} : (p^*,q^*) \notin [\widetilde{p},\infty) \times [\widetilde{q},\infty)\, \, \forall (\widetilde{p},\widetilde{q}) \in \widetilde{\cp}_{\textup{ALG-trct}}\setminus \{p^*,q^*\}  \}.
\]
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NovWoz08a}.

Having defined the tractability function as above, we will now examine a specific case where we always set $d=1$, i.e.,
\[
T(\varepsilon^{-1},1,\vp)= \max\{1,\varepsilon^{-\vp}\}\qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]
where in this case $\vp=p$ is a scalar.
